\emph{hp}-Finite Element Model of Poisson and Nernst-Planck System of Equations

### 2010-10-09 motivation.tex
Started results section, some images
Draft getting ready

\section{Motivation}
\begin{figure}
  \begin{centering}
      \includegraphics[scale=0.4]{comsol_conc_volt}
  \caption{\label{fig:comsol-conc-volt}Calculated concentration $C$ and voltage $\phi$
  in a 1D domain $\Omega\subset\mathbb{R}$.
  Dirichlet boundary conditions ($V_{\partial \Omega_1}=0\ V$
  and $V_{\partial \Omega_2}=4\ V$) were
	applied to the Poisson equation Eq.~\eqref{eq:poisson} and Neumann boundaries
	to the Nernst-Planck equation Eq.~\eqref{eq:nernst-planck}.}
  \end{centering}
\end{figure}
Fig.~\ref{fig:comsol-conc-volt} shows a solution for $C$ and $\phi$ in
the time dependent NPN system at $t=0.1\ s$ and $t=3.0\ s$. Solution
constants are shown in Table~\ref{Table:used-constants}.
It can be seen that the solution has two notable characteristics. 
Firstly, $\nabla C$ at $\partial \Omega_2$ is
moving in time, whereas $\nabla C$ at $\partial \Omega_1$ is very
sharp. For the large part of the domain $\Omega$ $\nabla C=0$.
At the same time, $\phi$ is a rather smooth fucntion in $\Omega$.
This clearly makes the choice of mesh for the time dependent NPN
system difficult. By creating a too fine mesh, the problem gets
too big. By choosing a coarser mesh, the calculation error goes up.
Furthermore, from the shape of the solution in Fig.~\ref{fig:comsol-conc-volt}
can be seen that the polynomial degree of the elements in the middle
of the domain $\Omega$ and near the boundaries $\partial \Omega_1,\ \partial \Omega_2$
should be different --- namely higher degree near the boundaries. 
Overall shape of the solutions suggest that
using different meshes for variables $C$ and $\phi$ can be beneficial
in terms of resources and calculation time.

All the aforementioned difficulties can be solved by using
a time dependent adaptive multimesh \emph{hp}-FEM solver.
Automatic time dependent adaptivity in conjuction with \emph{hp}-FEM 
helps to limit the error of the calculations by choosing
a suitable mesh and polynomial degrees of the elements at each time step.
In this work Hermes2D~\cite{Hermes-project} was used to study
the problem size, error convergence and solution time of NPN system 
with different adaptivity algorithms.

@@

\subsection{Practical application}
IPMCs have been studied in past two decades in view of using them
as noiseless mechanoelectrical or electromechanical transducers.
The advantages of the IPMCs over other electroactive polymer actuators
are low voltage bending, high strains ($>1\%$), and ability to work
in wet environments. A typical IPMC consists of a thin sheet of polymer
(often Nafion or Teflon) which is sandwiched between noble
metal electrodes such as platinum or gold. When fabricated, the polymer 
membrane is saturated with certain solvent and ions, e.g water and $H^+$.
When a voltage is applied to the eletrodes, the counter ions start
migrating due to the imposed electric field. By dragging along the solvent,
the osmotic pressure difference near the electrodes occur which in turn
results in bending of the material (see Fig.~\ref{fig:conceptual}).
\begin{figure}
  \begin{centering}
    \subfloat[~]{\begin{centering}
      \includegraphics[scale=0.6]{IPMC}
      \par\end{centering}}~
    \subfloat[~]{\begin{centering}
      \includegraphics[scale=0.6]{IPMC_bent}
      \par\end{centering}}
  \par\end{centering}
  \caption{\label{fig:conceptual}Conceptual model of the actuation
 	of IPMC. Initial counter ion distribution (a) and
	the distribution and resulting bending after applying a voltage (b).}
\end{figure}
The derived and implemented model helps to predict the actuation of the
material. Furthermore, it is expected that the \emph{hp}-FEM implementation
results in a smaller problem size, thus likely allowing faster calculation
in both 2D and 3D.

>>>>

\section{Motivation}
\begin{figure}
  \begin{centering}
      \includegraphics[]{comsol_conc_volt}
  \caption{\label{fig:comsol-conc-volt}Calculated concentration $C$ and voltage $\phi$
  in a 1D domain $\Omega\subset\mathbb{R}$.
  Dirichlet boundary conditions ($V_{\partial \Omega_1}=0\ V$
  and $V_{\partial \Omega_2}=4\ V$) were
	applied to the Poisson equation Eq.~\eqref{eq:poisson} and Neumann boundaries
	to the Nernst-Planck equation Eq.~\eqref{eq:nernst-planck}.}
  \end{centering}
\end{figure}
Fig.~\ref{fig:comsol-conc-volt} shows a solution for $C$ and $\phi$ in
the time dependent NPN system at $t=0.1\ s$ and $t=3.0\ s$. Solution
constants are shown in Table~\ref{Table:used-constants}.
It can be seen that the solution has two notable characteristics. 
Firstly, $\nabla C$ at $\partial \Omega_2$ is
moving in time, whereas $\nabla C$ at $\partial \Omega_1$ is very
large. For the most part of the domain $\Omega$ $\nabla C=0$.
At the same time, $\phi$ is a rather smooth fucntion in $\Omega$.
This clearly makes the choice of the mesh for the time dependent NPN
system difficult. By choosing a too fine mesh, the problem gets
too big. By choosing a coarse mesh, the calculation error might
get too large.
Furthermore, from the shape of the solution in Fig.~\ref{fig:comsol-conc-volt}
can be seen that the polynomial degree of the elements in the middle
of the domain $\Omega$ and near the boundaries $\partial \Omega_1,\ \partial \Omega_2$
should be different --- namely higher degree elements near the boundaries
and lower degree element in the middle of domain. 
Overall shape of the solutions also suggest that
using different meshes for variables $C$ and $\phi$ can be beneficial
in terms of number of degrees of freedom and calculation time.

All the aforementioned concerns can be solved by using
a time dependent adaptive multimesh \emph{hp}-FEM solver.
Automatic time dependent adaptivity in conjuction with \emph{hp}-FEM 
helps to limit the error of the calculations by automatically choosing
a suitable mesh and polynomial degrees of the elements at each time step.
In this work Hermes~\cite{Hermes-project} was used to implement
the PNP system and study
the problem size, error convergence and solution time 
with different refinement modes.

@@

\subsection{Practical application}
IPMCs have been studied in past two decades in view of using them
as noiseless mechanoelectrical or electromechanical transducers.
The advantages of the IPMCs over other electroactive polymer actuators
are low voltage bending, high strains ($>1\%$), and ability to work
in wet environments. A typical IPMC consists of a thin sheet of polymer
(often Nafion or Teflon) which is sandwiched between noble
metal electrodes such as platinum or gold. When fabricated, the polymer 
membrane is saturated with certain solvent and ions, e.g water and $H^+$.
When a voltage is applied to the eletrodes, the counter ions start
migrating due to the imposed electric field. By dragging along the solvent,
the osmotic pressure difference near the electrodes occur which in turn
results in bending of the material (see Fig.~\ref{fig:conceptual}).
\begin{figure}
  \begin{centering}
    \subfloat[~]{\begin{centering}
      \includegraphics[scale=0.6]{IPMC}
      \par\end{centering}}~
    \subfloat[~]{\begin{centering}
      \includegraphics[scale=0.6]{IPMC_bent}
      \par\end{centering}}
  \par\end{centering}
  \caption{\label{fig:conceptual}Conceptual model of the actuation
 	of IPMC. Initial counter ion distribution (a) and
	the distribution and resulting bending after applying a voltage (b).}
\end{figure}
The derived and implemented model helps to predict the actuation of the
material. Furthermore, it is expected that the \emph{hp}-FEM implementation
results in a smaller problem size, thus likely allowing faster calculation
in both 2D and 3D domains.

### 2010-10-09 pugal-esco2010.tex
Started results section, some images
Finished the results section
Draft getting ready

\begin{abstract}
 In this paper we present a \emph{hp}-finite element
model (\emph{hp}-FEM) of Nernst-Planck and Poisson equation system.
The model is implemented in Hermes2D which is a
space- and space-time adaptive \emph{hp}-FEM solver. The time dependent
adaptivity is used to control the error of the solution. Full mathematical
derivation of the weak formulation of the system of equations and
the solution comparison with a nonadaptive conventional FEM is presented.
Furthermore, we extend the discussion on Hermes3D as a modeling tool that can
help in studying and improving the IPMC materials.
\end{abstract}

@@

\input{intro}
\input{motivation}
\input{model}
\section{Results}
Some convergence grahps

\section{Conclusion, future works}

>>>>

\begin{abstract}
 In this paper we present a \emph{hp}-finite element
model (\emph{hp}-FEM) of Nernst-Planck and Poisson equation system.
The model is implemented in Hermes which is a
space- and space-time adaptive \emph{hp}-FEM solver. The time dependent
adaptivity is used to control the error of the solution. Full mathematical
derivation of the weak formulation of the system of equations and
the solution comparison with a nonadaptive conventional FEM is presented.
Furthermore, we extend the discussion on Hermes3D as a modeling tool that can
help in studying and improving the IPMC materials.
\end{abstract}

@@

\input{intro}
\input{motivation}
\input{model}
\input{results}
\input{conclusions}

\section*{Acknowledgments}
KJK thanks the financial support from the Office of Naval Research under Award N000140910218.

### 2010-10-09 results.tex
Draft getting ready

\begin{figure}
  \begin{centering}
  \includegraphics[width=.8\columnwidth]{mesh}
  \caption{\label{fig:mesh} Initial coarse mesh (a),
  	half refined mesh (b) and refined mesh (c). The coarse mesh
	and refined mesh were used in the initial calcualtions, the latter one
	in case of \emph{p}-adaptivity (including HP\_ANISO\_P). The half-refined mesh was
	used later to optimize \emph{hp}-adaptive mode solutions.}
  \end{centering}
\end{figure}
We ran the calculation for all the adaptivity types 
in both single-mesh and multi-mesh configurations. 
The following numerical results were recorded for each 
adaptivity type: converged relative error, cumulative CPU
time, and the problem size in terms of number
of degrees of freedoms (NDOFs) at each time step.  
Two types of initial meshes were used --- in case of only \emph{p}-adaptivity,
more refined mesh was used (see Fig.~\ref{fig:mesh}~(b)) to ensure
the error convergence.
When also the element size refinement
was enabled (all \emph{h} adaptivity types), very coarse initial mesh
was used (see Fig.~\ref{fig:mesh}~(a)) to let the adaptivity
algorithm find the most optimal mesh. 
In both case, the initial mesh was loaded at each time step of the
calculation.

\begin{figure}
  \begin{centering}
  \includegraphics[width=.75\columnwidth]{cphi}
  \caption{\label{fig:cphi} Concentration $C$
  and voltage $\phi$ at two different time steps
  (HP\_ANISO adaptivity was used). This is a 2D solution shown
  as in 3D, where the height indicates the value.}
  \end{centering}
\end{figure}
Example of the solution at $t=0.7\ s$ and $t=3.0\ s$ 
with adaptivity type HP\_ANISO is shown
in Fig.~\ref{fig:cphi}. The time $t=0.7\ s$ was chosen because
by that time, some ionic migration has already taken place, i.e.
concentration gradient near the $\partial\Omega_1$ and
$\partial\Omega_3$ has formed. The automatic mesh refinements
at different time steps are clearly visible --- especially
near the top boundary, where the concentration gradient is
moving in time (as seen in Fig.~\ref{fig:comsol-conc-volt}).

The following subsections provide a detailed comparison of
different adaptivity types and estimate the most suitable
one for the given problem.

\subsection{Optimal adaptivity types}

\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{singlemulti_dof}
  \caption{\label{fig:singlemultidof} NDOFs in case 
  of single-mesh and multi-mesh solutions with HP\_ISO
  and HP\_ANISO adaptivities. (Notice log Y scale)}
  \end{centering}
\end{figure}
\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{singlemulti_cpu}
  \caption{\label{fig:singlemulticpu} CPU times in case
  of single-mesh and multi-mesh solutions with HP\_ISO
  and HP\_ANISO adaptivities. (Notice log Y scale)}
  \end{centering}
\end{figure}
\begin{figure}
  \begin{centering}
  \includegraphics[width=.75\columnwidth]{poly}
  \caption{\label{fig:poly} Polynomial degree space
  for $C$ and $\phi$ at $t=0.7\ s$ The color indicates
  the maximum polynomial degree of the corresponding element.}
  \end{centering}
\end{figure}

Running the simulation with different adaptivity
types and meshes showed that the multi-mesh configuration generally results in
smaller problem size, faster calculation, and better or similar error convergence
compared to the single-mesh configuration.
This is well illustrated in Fig.~\ref{fig:singlemultidof} 
and Fig.~\ref{fig:singlemulticpu}.
This can be well understood from Fig.~\ref{fig:comsol-conc-volt} --- in the boundary
regions $\partial \Omega_1$ and $\partial\Omega_3$ the concentration gradient
is greater than voltage gradient, namely $\nabla C >> \nabla \phi$. Therefore
refining the mesh for both variables is not reasonable. For instance,
the solution space with corresponding mesh in case of
HP\_ANISO at $t=0.7\ s$ is shown in Fig.~\ref{fig:cphi}. The corresponding polynomial
degree space is shown in Fig.~\ref{fig:poly}. Notice that the adaptive algorithm
has increased the maximum polynomial degree for $C$ space to 7, however,
at the same time, maximum polynomial degree for $\phi$ is one. Furthermore,
the mesh is more refined for $C$.
For most cases using the multi-mesh results in similar or better CPU time, i.e.
multi-mesh takes less computational resources. Only excpetion was HP\_ANISO\_H 
adaptivity type for which the single mesh configuration resulted slightly faster
calculation time. 
Based on the results only multi-mesh configurations will be considered
in the following comparisons.

\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoaniso_dof}
  \caption{\label{fig:isoanisodof} NDOFs in case 
  of multi-mesh solutions with H\_ISO, H\_ANISO,
  HP\_ISO, HP\_ANISO, and single-mesh solution with HP\_ANISO\_H
  adaptivities. (Notice log Y scale)}
  \end{centering}
\end{figure}
\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoaniso_cpu}
  \caption{\label{fig:isoanisocpu} CPU times in case 
  of multi-mesh solutions with H\_ISO, H\_ANISO,
  HP\_ISO, HP\_ANISO, and single-mesh solution with HP\_ANISO\_H
  adaptivities. (notice log Y scale)}
  \end{centering}
\end{figure}
\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoanisop_dof}
  \caption{\label{fig:isoanisopdof} NDOFs in case 
  of multi-mesh solutions with P\_ISO, P\_ANISO, and
  HP\_ANISO\_P adaptivities.}
  \end{centering}
\end{figure}
\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoanisop_cpu}
  \caption{\label{fig:isoanisopcpu} CPU times in case 
  of multi-mesh solutions with P\_ISO, P\_ANISO, and
  HP\_ANISO\_P adaptivities.}
  \end{centering}
\end{figure}
To narrow down the list of adaptivity modes for the given problem, first the
isotropic and anisotropic adaptivities were compared. Here \emph{p}-adaptivity
modes are compared separately as they use more refined mesh.
Fig.~\ref{fig:isoanisodof} and Fig.~\ref{fig:isoanisocpu} show the comparison
of H\_ISO, H\_ANISO, HP\_ISO, HP\_ANISO, and HP\_ANISO\_H modes in terms of CPU time
and problem size.
Fig.~\ref{fig:isoanisopdof} and Fig.~\ref{fig:isoanisopcpu} show the similar
comparsion for the P\_ISO, P\_ANISO, and HP\_ANISO\_P modes.
Here we see that the anisotropic adaptivities result in a reasonable problem
size and the problems solve within a reasonable calculation time. It is interesting
to note that HP\_ANISO results in the smallest problem size. Also, in
\emph{p}-adaptivity group, HP\_ANISO\_P results in the smallest problem size
at each time step, whereas P\_ISO and P\_ANISO have a very largy problem size
during the first time steps of the solution.
Here the term ``reasonable problem size''
means that the number of degrees of freedom in time converges
to so that $N_{dof}<500$, and the term ``reasonable calculation time''
means that the calculation (step $\tau=0.01\ s$, physical
time $t_{end}=3.0\ s$) time $t$ on a givem system was $t<500\ s$.
Although these parameters are empirical, they serve as an upper limit, given
that the most adaptivity modes gave significantly smaller results,
e.g. $t<<500\ s$ and $N_{dof} << 500$.

\subsection{Quantitative analysis of different adaptivity types}

Based on the results of the previous subsection, H\_ANISO, HP\_ANISO,
and HP\_ANISO\_H from the \emph{hp/h} adaptivity group and HP\_ANISO\_P
from \emph{p} adaptivity group will be compared
in terms of problem size and cumulative CPU time. 
In all of the cases, the relative 
error at each time step remained below
the threshold which were set to $e_{th}=0.5\%$ between the coarse mesh
and fine mesh solutions, therefore the error vs. time plot will not be considered.

\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{cpu}
  \caption{\label{fig:cpu} Comparative CPU time for different adaptivity modes.}
  \end{centering}
\end{figure}

\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{dof}
  \caption{\label{fig:dof} Comparative NDOFs at each time step for 
  different adaptivity modes.}
  \end{centering}
\end{figure}

Fig.~\ref{fig:cpu} shows the cumulative CPU time for different adaptivity
modes at each time step. All the calculations were done on the same computer.
Here we see that HP\_ANISO\_H and HP\_ANISO require the most
resources which can be understood from the fact that these 
adaptivity modes have the largest number of 
candidates (see XXX) from which the refinement methode is chosen.
Fig.~\ref{fig:dof} shows the NDOFs at each time step for different adaptivity modes.
There we see that the HP\_ANISO results in the 
smallest problem size --- $N_{dof} \approx 125$. 
All the other adaptivity modes result in the 
problem size of approximately ($N_{dof} \approx 250$),

\subsection{Optimizations}

@@

So far we have seen that HP\_ANISO results in the smallest problem size, but requires
quite a lot CPU time. At the same time, HP\_ANISO\_P is the fastest, whereas
it does not result in that small problem. When it comes to a large domain 
or 3D modeling the problem
size becomes the most important factor. Therefore, we consider HP\_ANISO the most
suitable adaptivity type to the given problem. Thus some ways to optimize the time
factor will be considered. The desirable output would be HP\_ANISO problem size
close to HP\_ANISO\_P CPU time. One way to optimize the problem
is to choose somewhat more refined initial mesh. Other way one could think
to optimize the problem would be to change
the refinement frequency in a time dependent way. Recall that up to this point,
the initial mesh has been loaded in the beginning of each time step. True,
by employing the optimizations, we already must know something about the problem
and its solution beforehand. However, it could still be practial when solving a real problem
in a large domain.

The problem size and CPU time with HP\_ANISO
and HP\_ANISO\_P adaptivities on more refined initial mesh (see Fig.~\ref{fig:mesh}~(c))
compared to the coarse initial mesh (Fig.~\ref{fig:mesh}~(a)) and 
HP\_ANISO\_P solution is shown in Fig.~\ref{fig:refined-cpu} and Fig.~\ref{fig:refined-dof}.
It can be seen that by using initially more refined mesh, et problem solving time
can be reduced in case of HP\_ANISO, at the same time, the problem size increases.
In this situation, HP\_ANISO\_P and HP\_ANISO perform equally well.

\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{unreffreq_dof}
  \caption{\label{fig:unreffreq-dof} NDOFs at each time step for
  HP\_ANISO and HP\_ANISO\_P with mesh unrefinement at each time step and at over each
  	time step after first $0.5\ s$ physical solution time.}
  \end{centering}
\end{figure}
The next proposed optimization involved changing the unrefinement frequency.
It is known that the concentration gradient $\partial C$ changes the most in the initial phase
of the calculation, therefore, the unrefinement after each time step was performed
until $t=0.5\ s$ (physical time). After that, the
unrefinement was performed in $0.10\ s$ interval.
It appears that this type
of optimization does not result in a stable solution, i.e. the problem size 
does not remain steady, but starts to oscillate depending on the unrefinement
frequency. This is demonstrated in 
Fig.~\ref{fig:unreffreq-dof}. Therefore this kind of optimization did not
prove to be efficient.

\subsection{More general results}
\begin{figure}
  \begin{centering}
  \includegraphics[width=.75\columnwidth]{cphiorders}
  \caption{\label{fig:cphi-orders} Solutions $C$ and $\phi$
  and corresponding polynomial degrees of the elements at
  $t=0.7\ s$. HP\_ANISO adaptivity type was used. The height
  in the solution graphs indicates the value.}
  \end{centering}
\end{figure}

Based on the results, cation concentration and voltage was calculated
for different boundary conditions.
For instance, when voltage $\phi_{\Omega_1}=0.5\frac{x}{width_{\Omega_1}}+0.5$, the
concentration gradient $\nabla C$ and the voltage gradient $\nabla \phi$ are no
longer effectively 1D.
The calculated $C$ and $\phi$ in $\Omega$ and corresponding meshes and polynomial
degrees of the elements are shown in Fig.~\ref{fig:cphi-orders}.
HP\_ANISO adaptivity type was used. Notice that the solution
is different to the one in Fig.~\ref{fig:cphi} and the adapted mesh and the
polynomial degrees are also more complicated than in Fig.~\ref{fig:poly}.
It must be also noted that in case of non uniform boundary conditions which
results in 2D problem, refined initial mesh is more efficient to use.

>>>>

\begin{figure}
  \begin{centering}
  \includegraphics[width=.8\columnwidth]{mesh}
  \caption{\label{fig:mesh} Initial coarse mesh (a),
  	half refined mesh (b) and refined mesh (c). The coarse mesh
	and refined mesh were used in the initial calculations, the latter one
	in case of \emph{p}-adaptivity (including HP\_ANISO\_P). The half-refined mesh was
	used later to optimize the \emph{hp}-adaptive refinement solutions.}
  \end{centering}
\end{figure}
Calculation for all the refinement modes were performed
in both single-mesh and multi-mesh configurations. 
The following numerical results were recorded for each 
refinement mode: converged relative error, cumulative CPU
time, and the problem size in terms of number
of degrees of freedoms (NDOFs) at each time step. 
Two types of initial meshes were used --- in case of only \emph{p}-adaptivity,
more refined mesh was used (Fig.~\ref{fig:mesh}~(c)) to ensure
the error convergence.
When the element size refinement
was also enabled (all \emph{h}/\emph{hp} refinement modes), very coarse initial mesh
was used (see Fig.~\ref{fig:mesh}~(a)) to let the adaptivity
algorithm find the most optimal mesh. So coarse initial mesh might not be
suitable for all the practical applications, as will be demonstrated in the
end of this section, however, it provides a good insight into the
adaptivity performance of Hermes.
In both cases, the initial mesh was loaded at each time step of the
calculation.

\begin{figure}
  \begin{centering}
  \includegraphics[width=.75\columnwidth]{cphi}
  \caption{\label{fig:cphi} Concentration $C$
  and voltage $\phi$ at two different time steps
  (HP\_ANISO adaptivity was used). This is a 2D solution shown
  as in 3D, where the height indicates the values of $C$ and $\phi$.}
  \end{centering}
\end{figure}
Example of the solution at $t=0.7\ s$ and $t=3.0\ s$ 
with refinement mode HP\_ANISO is shown
in Fig.~\ref{fig:cphi}. The time $t=0.7\ s$ was chosen because
by that time, some ionic migration has already taken place, i.e.
concentration gradient near the $\partial\Omega_1$ and
$\partial\Omega_3$ has formed. The automatic mesh refinements
at different time steps are clearly visible in the figure --- especially
near the top boundary, where the concentration gradient is
moving in time (as was also seen in Fig.~\ref{fig:comsol-conc-volt}).

The following subsections provide a detailed comparison of
the different refinement modes and try to estimate the most suitable
one for the given problem.

\subsection{Optimal refinement modes}

\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{singlemulti_dof}
  \caption{\label{fig:singlemultidof} NDOFs in case 
  of single-mesh and multi-mesh solutions with HP\_ISO
  and HP\_ANISO refinements. (Notice log Y scale)}
  \end{centering}
\end{figure}
\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{singlemulti_cpu}
  \caption{\label{fig:singlemulticpu} CPU times in case
  of single-mesh and multi-mesh solutions with HP\_ISO
  and HP\_ANISO refinements. (Notice log Y scale)}
  \end{centering}
\end{figure}
\begin{figure}
  \begin{centering}
  \includegraphics[width=.75\columnwidth]{poly}
  \caption{\label{fig:poly} Polynomial degree space
  for $C$ and $\phi$ at $t=0.7\ s$ The color indicates
  the maximum polynomial degree of the corresponding element.}
  \end{centering}
\end{figure}

Running the simulation with different refinement modes 
and meshes showed that the multi-mesh configuration generally results in
a smaller problem, faster calculation, and better or similar error convergence
compared to the single-mesh configuration.
This is well illustrated in Fig.~\ref{fig:singlemultidof} 
and Fig.~\ref{fig:singlemulticpu}.
The result can be understood from Fig.~\ref{fig:comsol-conc-volt} --- in the boundary
regions $\partial \Omega_1$ and $\partial\Omega_3$ the concentration gradient
is greater than voltage gradient, namely $\nabla C >> \nabla \phi$. Therefore
refining the mesh for both variables is not reasonable in terms of
number of degrees of freedom. For instance,
the solution space with corresponding mesh in case of
HP\_ANISO at $t=0.7\ s$ is shown in Fig.~\ref{fig:cphi}. The corresponding polynomial
degree space is shown in Fig.~\ref{fig:poly}. Notice that the adaptive algorithm
has increased the maximum polynomial degree for $C$ space to 7, however,
at the same time, the maximum polynomial degree for the $\phi$ space is one. Furthermore,
the mesh is significantly more refined for $C$.
For most cases using the multi-mesh results in similar or better CPU time, i.e.
multi-mesh takes less computational resources. Only excpetion was HP\_ANISO\_H 
refinement mode for which the single mesh configuration resulted slightly faster
calculation time. 
Based on the results, only multi-mesh configurations will be considered
in the following comparisons.

\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoaniso_dof}
  \caption{\label{fig:isoanisodof} NDOFs in case 
  of multi-mesh solutions with H\_ISO, H\_ANISO,
  HP\_ISO, HP\_ANISO, and single-mesh solution with HP\_ANISO\_H
  refinement modes. (Notice log Y scale)}
  \end{centering}
\end{figure}
\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoaniso_cpu}
  \caption{\label{fig:isoanisocpu} CPU times in case 
  of multi-mesh solutions with H\_ISO, H\_ANISO,
  HP\_ISO, HP\_ANISO, and single-mesh solution with HP\_ANISO\_H
  refinement modes. (notice log Y scale)}
  \end{centering}
\end{figure}
\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoanisop_dof}
  \caption{\label{fig:isoanisopdof} NDOFs in case 
  of multi-mesh solutions with P\_ISO, P\_ANISO, and
  HP\_ANISO\_P refinement modes.}
  \end{centering}
\end{figure}
\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoanisop_cpu}
  \caption{\label{fig:isoanisopcpu} CPU times in case 
  of multi-mesh solutions with P\_ISO, P\_ANISO, and
  HP\_ANISO\_P refinement modes.}
  \end{centering}
\end{figure}
To narrow down the list of refinement modes for the given problem, first the
isotropic and anisotropic adaptivities were compared. The \emph{p}-adaptivity
modes were compared separately as they used more refined mesh (Fig.~\ref{fig:mesh}~(c)).
Fig.~\ref{fig:isoanisodof} and Fig.~\ref{fig:isoanisocpu} show the comparison
of H\_ISO, H\_ANISO, HP\_ISO, HP\_ANISO, and HP\_ANISO\_H modes in terms of CPU time
and problem size.
Fig.~\ref{fig:isoanisopdof} and Fig.~\ref{fig:isoanisopcpu} show the similar
comparsion for the P\_ISO, P\_ANISO, and HP\_ANISO\_P modes.
It can be clearly seen that the anisotropic refinement modes result in a reasonable problem
size and the problems solve within a reasonable calculation time. It is interesting
to note that HP\_ANISO results in the smallest problem size. Also, in
\emph{p}-adaptivity group, HP\_ANISO\_P results in the smallest problem size
at each time step, whereas P\_ISO and P\_ANISO have a very largy problem size
during the first time steps of the solution.
Here the term ``reasonable problem size''
means that the number of degrees of freedom in time converges
to so that $N_{dof}<500$, and the term ``reasonable calculation time''
means that the calculation (step $\tau=0.01\ s$, physical
time $t_{end}=3.0\ s$) time $t$ on a givem system was $t<500\ s$.
Although these parameters are empirical, they serve as an upper limit, given
that the most refinement modes give significantly smaller results:
$t<<500\ s$ and $N_{dof} << 500$.

\subsection{Quantitative analysis of the refinement modes}

Based on the results in the previous subsection, H\_ANISO, HP\_ANISO,
and HP\_ANISO\_H from the \emph{hp/h}-adaptivity group and HP\_ANISO\_P
from \emph{p}-adaptivity group will be compared
in terms of problem size and cumulative CPU time. 
In all of the cases, the relative 
error at each time step remained below
the threshold which were set to $e_{th}=0.5\%$ between the coarse mesh
and fine mesh solutions, therefore the error-time plot will not be considered.

\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{cpu}
  \caption{\label{fig:cpu} Comparative CPU time for different refinement modes.}
  \end{centering}
\end{figure}

\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{dof}
  \caption{\label{fig:dof} Comparative NDOFs at each time step for 
  different refinement modes.}
  \end{centering}
\end{figure}

Fig.~\ref{fig:cpu} shows the cumulative CPU time for different refinement 
modes at each time step. All the calculations were done on the same computer.
Here we see that HP\_ANISO\_H and HP\_ANISO require the most
resources which can be understood from the fact that these 
refinement modes have the largest number of 
candidates (see Section.~\ref{sec:model}) from which the refinement method is chosen.
At the same time, HP\_ANISO\_P is the fastest among all the other refinement modes.
Fig.~\ref{fig:dof} shows the NDOFs at each time step.
It can be seen that the HP\_ANISO results in the 
smallest problem size --- $N_{dof} \approx 125$. 
All the other refinement modes result in a 
problem size of approximately ($N_{dof} \approx 250$). So we have
two notable refinement modes --- HP\_ANISO\_P because of the fast
solution, and HP\_ANISO because of the small problem size. In the following
subsection, optimization of those two modes will be considered.

\subsection{Optimizations of HP\_ANISO and HP\_ANISO\_P refinement modes}

@@

So far we have seen that HP\_ANISO results in the smallest problem size, but requires
quite a lot CPU time compared to HP\_ANISO\_P that is the fastest.
When it comes to a large domain or 3D modeling the problem
size becomes the most important factor. Therefore, we consider HP\_ANISO the most
suitable refinement mode to the given problem. Thus some ways to optimize the time
factor will be considered. The desirable output would be HP\_ANISO problem size
close to HP\_ANISO\_P CPU time. One way to optimize the problem
is to choose somewhat more refined initial mesh. Other way one could think
to optimize the problem would be to change
the refinement frequency during the solving process. Recall that up to this point,
the initial mesh has been loaded in the beginning of each time step. True,
by employing the optimizations, we already must know something about the problem
and its solution beforehand. However, it could still be practial when solving a real problem
in a large domain.

The problem size and CPU time with HP\_ANISO
and HP\_ANISO\_P adaptivities on more refined initial mesh (see Fig.~\ref{fig:mesh}~(c))
compared to the coarse initial mesh (Fig.~\ref{fig:mesh}~(a)) and 
HP\_ANISO\_P solution is shown in Fig.~\ref{fig:refined-cpu} and Fig.~\ref{fig:refined-dof}.
By using initially more refined mesh, the problem solving time
can be reduced in case of HP\_ANISO, at the same time, the problem size increases.
In this situation, HP\_ANISO\_P and HP\_ANISO perform equally well.

\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{unreffreq_dof}
  \caption{\label{fig:unreffreq-dof} NDOFs at each time step for
  HP\_ANISO and HP\_ANISO\_P with mesh unrefinement at each time step and at over each
  	time step after first $0.5\ s$ physical solution time.}
  \end{centering}
\end{figure}
The next proposed optimization involved changing the unrefinement frequency.
It is known that the concentration gradient $\nabla C$ changes the most in the initial phase
of the calculation, therefore, the unrefinement after each time step was performed
until $t=0.5\ s$ (physical time). After that, the
unrefinement was performed in $\Delta t = 0.10\ s$ interval.
However, this optimization did not result in a stable solution, i.e. the problem size 
does not remain steady, but starts to oscillate depending on the unrefinement
frequency. This is shown in 
Fig.~\ref{fig:unreffreq-dof}.

Therefore varying the unrefinement frequency will
not likely result in desired results in real applications for given system of equation.
At the same time, by varying a mesh size, optimal initial mesh could be found
for both HP\_ANISO and HP\_ANISO\_P refinement modes.

\subsection{More general results}
\begin{figure}
  \begin{centering}
  \includegraphics[width=.75\columnwidth]{cphiorders}
  \caption{\label{fig:cphi-orders} Solutions $C$ and $\phi$
  and corresponding polynomial degrees of the elements at
  $t=0.7\ s$. HP\_ANISO refinement mode was used. The height
  in the solution graphs indicates the value.}
  \end{centering}
\end{figure}

Based on the results, cation concentration and voltage was calculated
for different boundary conditions.
For instance, when voltage is applied as follows
\begin{equation}
  \phi_{\Omega_1}=0.5\frac{x}{width_{\Omega_1}}+0.5,
\end{equation}
the concentration gradient $\nabla C$ and the voltage gradient $\nabla \phi$ are no
longer effectively 1D.
The calculated $C$ and $\phi$ in $\Omega$ and corresponding meshes and polynomial
degrees of the elements are shown in Fig.~\ref{fig:cphi-orders}.
HP\_ANISO refinement mode was used. Notice that the solution
is different to the one in Fig.~\ref{fig:cphi} and the adapted mesh and the
polynomial degrees are also more complicated than in Fig.~\ref{fig:poly}.
It must be noted that in case of non uniform boundary conditions which
results in 2D problem, refined initial mesh was more efficient to use.

\emph{
PAVEL: Should I try to include the similar solution from comsol and do
some comparison here? I got this idea just before commiting the final
draft to you?}
